Towards a Neural Co-Processor Which Restores Dexterity After Stroke: Modeling a Proof-of-Concept

### 2021-12-15 paper/article.tex
Paper progress

TODO: describe stopping conditions

@@

\begin{table}[]
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{Experiment}  & \textbf{\begin{tabular}[c]{@{}l@{}}Lesioned\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task val\\ loss\end{tabular}} & \textbf{Pct recov} & \textbf{\begin{tabular}[c]{@{}l@{}}Pct recov\\ val\end{tabular}} \\ \hline
AIP No-coadapt & 0.004507                                                          & 0.004599                                                         & 0.004658                                                             & -2.32\%            & -3.82\%                                                          \\ \hline
AIP Coadapt    & 0.004507                                                          & 0.001575                                                         & 0.001502                                                             & 74.41\%            & 76.26\%                                                          \\ \hline
M1 No-coadapt  & 0.021136                                                          &                                                                  &                                                                      & 102.75\%           & 102.75\%                                                         \\ \hline
M1 Coadapt     & 0.021136                                                          & 0.004772                                                         & 0.004815                                                             & 79.55\%            & 79.35\%                                                          \\ \hline
Con No-coadapt & 0.020719                                                          & 0.002584                                                         & 0.002583                                                             & 89.99\%            & 89.99\%                                                          \\ \hline
Con Coadapt    & 0.020719                                                          & 0.002583                                                         & 0.002405                                                             & 89.99\%            & 90.88\%                                                          \\ \hline
Pre-recovery*  & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
\end{tabular}
\caption{\label{tab:results}Losses and recovery. *Pre-recovery lesioned loss and percent recoveries based on the post-recovery values.}
\end{table}

@@

\subsection{Experiment 1: AIP 75\% lesion}
This experiment demonstrates the ability for the co-processor to succeed after the
patient has lost some of the neural machinery responsible for representing the task.
In the non-coadaptive version of this experiment, the loss of AIP neurons involved
in encoding the visual inputs results in the simulated brain being unable to
condition its grasp for the object shape, though reaching largely still succeeds.
Likewise, the co-processor doesn't have sufficient information to improve performance.
In this case, the co-processor is effectively unable to improve on existing performance,
for lack of that information.

@@

\subsection{Experiment 2: M1 75\% lesion}
TODO: after final results
* M1 + coadaptation shows we can adapt to changes in the output mappings
* Initially, performance improves orders of magnitude more quickly than later. 50-75\% recovery during this phase. This appears to be the co-processor learning the task structure (e.g. not to move the hand until after the hold signal is lifted).
* Later: fine tuning and object class separation. Learning task structure.

\subsection{Experiment 3: F5 and M1 connectivity lesion}
As in the other cases, the co-processor's loss in this experiment improved
quickly, initially, but takes a long time to refine. During that refinement,
object differentiation occurs. Note that in this experiment, no information
can propagate forward to the M1 module, since we have completely severed the
input connections to that module. It is the sole responsibility of the
co-processor to perform that forward-propagation. As a result, co-adaptation
only affects the M1 module. The co-processor's learning algorithm successfully
adapted to this case, where the observed brain activity was stationary, but
the mapping between stimulation and task performance was not.

@@

The co-processor in this scenario successfully improved task performance.
Note that the object separation class separation metric $S$ returned to
a healthy value towards the end of the training, indicating that the
learned solution did not simply refine the stereotyped grasp. That in
turn demonstrates the co-processor's ability to propagate information
``around'' the simulation lesion.

>>>>

\subsection{Stopping condition}
For the purpose of these experiments, we train the co-processor until one of two
conditions is satisfied:

\begin{itemize}
	\item The percent change in task loss between two consecutive 500 epoch windows is below a certain threshold: 0.01\%. This
	      stopping criteria ensures we stop training when the economy of additional training is low.
	\item The epoch count is greater than 250k. This criteria provides a backstop against very slow training running without end.
\end{itemize}

@@

\begin{table}[]
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{Experiment}  & \textbf{\begin{tabular}[c]{@{}l@{}}Lesioned\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task val\\ loss\end{tabular}} & \textbf{Pct recov} & \textbf{\begin{tabular}[c]{@{}l@{}}Pct recov\\ val\end{tabular}} \\ \hline
AIP No-coadapt & 0.004507                                                          & 0.004599                                                         & 0.004658                                                             & -2.32\%            & -3.82\%                                                          \\ \hline
AIP Coadapt    & 0.004507                                                          & 0.001575                                                         & 0.001502                                                             & 74.41\%            & 76.26\%                                                          \\ \hline
M1 No-coadapt  & 0.021136                                                          & 0.005765                                                         & 0.005604                                                             & 74.73\%            & 75.51\%                                                          \\ \hline
M1 Coadapt     & 0.021136                                                          & 0.004772                                                         & 0.004815                                                             & 79.55\%            & 79.35\%                                                          \\ \hline
Con No-coadapt & 0.020719                                                          & 0.002584                                                         & 0.002583                                                             & 89.99\%            & 89.99\%                                                          \\ \hline
Con Coadapt    & 0.020719                                                          & 0.002583                                                         & 0.002405                                                             & 89.99\%            & 90.88\%                                                          \\ \hline
Pre-recovery*  & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
\end{tabular}
\caption{\label{tab:results}Losses and recovery. *Pre-recovery lesioned loss and percent recoveries based on the post-recovery values.}
\end{table}

@@

\subsection{Experiment 1: AIP 75\% lesion}
This experiment demonstrates the ability for the co-processor to succeed after the
patient has lost some of the neural machinery responsible for representing the objects' shapes.
In the non-coadaptive version of this experiment, the loss of AIP neurons involved
in encoding the visual inputs results in the simulated brain being unable to
condition its grasp for the object shape, though reaching largely still succeeds.
Likewise, the co-processor doesn't have sufficient information to improve performance,
for lack of that information.

@@

\subsection{Experiment 2: M1 75\% lesion}
In this experiment, 75\% of the output module of the simulated network has been
lesioned. In the coadaptive case, the co-processor needs to adapt to
non-stationarity in the mapping between stimulation and task performance.
In the non-coadaptive case, training quickly plateaus, and does not improve
thereafter. That suggests we have lost some of our degrees of freedom to control
the activations in the output layer, resulting in a reduced ability to recreate
the target grasps.

In the co-adapative case, performance likewise hits an inflexion point, but
continues to improve slowly.  TODO: finish once we have the faster co-adaptation
rate tested.

\subsection{Experiment 3: F5 and M1 connectivity lesion}
As in the other cases, the co-processor's loss in this experiment improves
quickly, initially, but taskes much longer to refine. During that refinement,
object differentiation occurs. Note that in this experiment, no information
can propagate forward to the M1 module, since we have completely severed the
input connections to that module. It is the sole responsibility of the
co-processor to perform that forward-propagation.

In this case, co-adaptation only affects the M1 module, since we have nothing
to use as a learning signal: there is no backpropagation to the AIP and
F5 modules. As a result, this experiment primarily demonstrates that the
co-processor's learning algorithm is capable of adapting to non-stationarity
in the mapping between stimulation parameters and task performance. Contrast
that with the results of Experiment 1, where the non-stationarity affected
the mapping between the visual inputs and the brain observations.

@@

The co-processor in this scenario successfully improved task performance.
Note that the object class separation metric $S$ returned to
a healthy value towards the end of the training, indicating that the
learned solution did not simply refine the stereotyped grasp.
